Predictive Quantization for MIMO-OFDM SVD Precoders using Reservoir Computing Framework

### 2019-04-30 paper/main.tex

\subsection{Quantization error, BER and Achievable Rate Results}
\label{res}{}
To mitigate errors in tracking in the presence of randomly appearing drastic events in wireless channels, such as sudden interferences and dynamic environments, only 100 channel evolutions are considered for the simulations.
Averaging is performed over 10 independent channel realizations, to ensure that the simulation results hold in general for the chosen IEEE Pedestrian A channel profile. Therefore, the quantization error, BER and achievable rate results are averaged over 1000 channel instances, with 100 channel evolutions of 10 independent channel realizations.
\begin{figure}[h]
\centering
\includegraphics[width=0.45\textwidth]{images/qtizErr.pdf}
\caption{Quantization error in terms of chordal distance to the actual precoder $\bU_{\{9k\},t}$ observed at the receiver, from the quantized estimate $\tilde{\bU}_{\{9k\},t}$ at the transmitter, averaged over 8 fed back subcarriers ($\{9k, k \in \{0,1,\ldots,7\}\}$) and 10 independent channel realizations (normalized Doppler $f_DT_s=10^{-4}$).}
\label{qtiz_err}
\end{figure}
From Fig. \ref{qtiz_err} it is evident that the reservoir computing based predictive quantizer is able to reduce the quantization error to below what was obtained from the existing temporal correlation based predictive quantizer in \cite{6891198}. In addition, the quantization error is much smoother, which can be explained from the fact that the reservoir computing approach optimizes, $\mathcal{O}_{i,t}$ that captures the long term dynamics of the channel and hence, is able to correct for jittery variations.

We simulate the (uncoded QPSK) BER performance for precoders obtained after interpolation ($\tilde{\bU}_{i\neq\{9k\},t}$) and quantization ($\tilde{\bU}_{\{9k\},t}$), for all the 63 subcarriers.
Observe from Fig. \ref{BER} that reservoir computing framework is able to achieve substantial improvements in $\text{E}_\text{b}\text{N}_0$ levels at $\text{BER}\leq 10^{-3}$.
Particularly, for $\text{BER}=10^{-4}$, we observe around $5$ dB improvement.
In addition, substantial improvement is obtained with both the predictive quantization schemes over the independent quantization scheme with Lloyd codebook for precoders \cite{6678348}, using the same number of bits.

For simulating the achievable rate for the point-to-point MIMO system, we quantize the channel matrix singular values ($\Sigma_{i,t}$ obtained from SVD of $\bH_{i,t}$) using a 2 bit vector quantizer ($k$-means), and feed them back to the transmitter for fed back subcarriers ($\{9k, k \in \{0,1,\ldots,7\}\}$).
For the non-fed back subcarriers, they are interpolated by nai\"ve linear interpolation, as in \cite{Gupt1905:Predictive}.
Observe from Fig. \ref{achievrate} that the reservoir computing framework offers significant improvement in achievable rate for normalized Doppler $f_DT_s=10^{-4}$.
Since temporal variations for normalized Doppler $f_DT_s=10^{-2}, 10^{-1}$ are higher, the reservoir computing framework has to learn a faster-varying function, and hence the degradation in performance.
However, it still matches the performance of the predictive quantizer in \cite{6891198} for normalized Doppler $f_DT_s=10^{-2}$ and slightly outperforms slightly for $f_DT_s=10^{-1}$.
\begin{figure}[h]
\centering
\includegraphics[width=0.45\textwidth]{images/BER_res.pdf}
\caption{BER results after obtaining non fed back $\tilde{\bU}_{i\neq\{9k\},t}$ from $\tilde{\bU}_{\{9k\},t}$ via Cayley interpolation \cite{Gupt1905:Predictive}, averaged over 10 independent channel realizations, 100 channel evolutions each (normalized Doppler $f_DT_s=10^{-4}$).}
\label{BER}
\end{figure}

>>>>

\subsection{Quantization error, BER and Achievable Rate Results}
\label{res}{}
To mitigate errors in tracking in the presence of randomly appearing drastic events in wireless channels, such as sudden interferences and dynamic environments, only 100 channel evolutions are considered for the simulations.
Averaging is performed over 10 independent channel realizations, to ensure that the simulation results hold in general for the chosen IEEE Pedestrian A channel profile. Therefore, the quantization error, BER and achievable rate results are averaged over 1000 channel instances, with 100 channel evolutions of 10 independent channel realizations.
\begin{figure}[h]
\centering
\includegraphics[width=0.45\textwidth]{images/qtizErr.pdf}
\caption{Quantization error in terms of chordal distance to the actual precoder $\bU_{\{9k\},t}$ observed at the receiver, from the quantized estimate $\tilde{\bU}_{\{9k\},t}$ at the transmitter, averaged over 8 fed back subcarriers ($\{9k, k \in \{0,1,\ldots,7\}\}$) and 10 independent channel realizations (normalized Doppler $f_DT_s=10^{-4}$).}
\label{qtiz_err}
\end{figure}

From Fig. \ref{qtiz_err} it is evident that the reservoir computing based predictive quantizer is able to reduce the quantization error below what was obtained from the existing temporal correlations based predictive quantizer in \cite{6891198}. In addition, the quantization error is much smoother, which can be explained from the fact that the reservoir computing approach optimizes, $\mathcal{O}_{i,t}$ that captures the long term dynamics of the channel and hence, is able to correct for jittery variations.

We simulate the (uncoded QPSK) BER performance for precoders obtained after interpolation ($\tilde{\bU}_{i\neq\{9k\},t}$) and quantization ($\tilde{\bU}_{\{9k\},t}$), for all the 63 subcarriers. Observe from Fig. \ref{BER} that reservoir computing framework is able to achieve substantial improvements in $\text{E}_\text{b}\text{N}_0$ levels at $\text{BER}\leq 10^{-4}$. Particularly, observe, for $\text{BER}=10^{-5}$ we observe around $5$ dB improvement.

For simulating the achievable rate for the point-to-point MIMO system considered, we quantize the sigma values ($\Sigma_{i,t}$ obtained from SVD of $\bH_{i,t}$) using a 2 bit vector quantizer ($k$-means), and feed them back to the transmitter for fed back subcarriers ($\{9k, k \in \{0,1,\ldots,7\}\}$). For the non-fed back subcarriers, they are interpolated via nai\"ve linear interpolation method as in \cite{Gupt1905:Predictive}. Observe from Fig. \ref{achievrate} that the reservoir computing framework offers signicant improvement in achievable rate for $f_DT_s=10^{-4}$. Since the temporal variations for $f_DT_s=10^{-2}, 10^{-1}$ are higher, the reservoir computing framework has to learn a faster-varying function, and hence the degradation in performance. However, it still matches the performance of the predictive quantizer in \cite{6891198} for $f_DT_s=10^{-2}$ and performs slightly better for $f_DT_s=10^{-1}$.

\begin{figure}[h]
\centering
\includegraphics[width=0.45\textwidth]{images/BER_res.pdf}
\caption{BER results after obtaining non fed back $\tilde{\bU}_{i\neq\{9k\},t}$ from $\tilde{\bU}_{\{9k\},t}$ via Cayley interpolation \cite{Gupt1905:Predictive}, averaged over 10 independent channel realizations, 100 channel evolutions each (normalized Doppler $f_DT_s=10^{-4}$).}
\label{BER}
\end{figure}
